Short Paper

### 2023-04-05 manuscript/manuscript.qmd
Add appendices

Queen's, simply, produces an immense amount of waste that it is unable
to efficiently manage. The dozens of wards, offices, and kitchens,
serving the hundreds of patients and staff, generates considerable
infectious medical waste, mixed domestic and office waste, and kitchen
waste, which needs to be removed and disposed of daily. Some individual
wards implement waste separation, binning medical wastes separately from
general waste, however separation ends at the point of collection.
Queen's is not served by regular municipal waste collection. What
municipal collection that is done is ad-hoc, sporadic, and based on the
hospital's available financial resources. Rather, most hospital waste
(infectious and non-infectious) is gathered by grounds staff and burnt
at the hospital's incinerator, located at a central point within the
hospital campus [@kalina2020this]. However, the vast majority of the
deposited waste is not incinerated, as a new incinerator provided by the
Ministry of Health in late 2019, is only able to handle a small
percentage of the hospital's waste. Furthermore, the incinerator is
frequently inoperable, under repair, or without power,. Rather, the
waste is openly burned, in a constantly smouldering pile to the side of
the incinerator building, sending up the aforementioned clouds of smoke,
which choke the grounds (Figure 1). Innumerable other, smaller, fires,
scattered across the QECH campus (Figure 2) contribute their part, as
grounds and maintenance staff habitually burn leaves and other garden
refuse. Furthermore, caregivers who reside separately on hospital
grounds, cook for themselves throughout the day and night using biomass
(charcoal or wood), and burn their own domestic waste. As a result, air
quality at the hospital is a constant source of discussion, with
patients, staff, and caregivers struggling to cope, when just breathing
is, as the quote used in the title of this article, 'unbearable.'

@@

-   **The Blantyre Malaria Project (BMP)**, established by Professors
    Terrie Taylor and Malcolm Molyneux, has carried out clinical
    research and patient care in the area of paediatric malaria,
    specifically cerebral malaria, for more than 30 years. Its research
    infrastructure includes an administrative team, an inpatient
    research ward, an MRI centre, and an outpatient research clinic in
    Ndirande township and in other districts outside of Blantyre.

@@

-   When a patient arrives at QECH, she or he will usually arrive with a
    guardian: someone to cook for them, buy medicine, do their laundry,
    and help them bathe. The **Guardian Shelter** is a gender-separated
    concrete floor shelter for sleeping along with a cooking pavilion
    and toilet/shower block are maintained by a local NGO. Though
    simple, the facility is secure and offers one of the only
    free/accessible toilet facilities for visitors on the QECH campus.

@@

The PM monitoring device centres around a Raspberry Pi 3 Model B
single-board computer complete with an operating system and storage
space. The added advantage of storage space is that the data are not
lost if the internet connection breaks down. Because the Pi can only
read digital signals, we needed to include an analog-to-digital (ADC)
converter. The amount of dust is measured by a Nova PM SDS011 High
Precision Laser sensor. This sensor measures particles at 2.5 and 10
micrometres in diameter in mg/m3. A list of the specific hardware and
software components can be found in Appendix TODO.

@@

Due to localised variation (air movement in the immediate area) and the
density of the observations, the plotted 5-minute data obscure
persistent trends. The full set of 5-minute data are plotted and
presented in Appendix XA1 (TODO). However, the number of measured values
that exceeded the hazardous limit for both parameters are summarised in
Table 2.

@@

# Appendix

@@

>>>>

Queen's, simply, produces an immense amount of waste that it is unable
to efficiently manage. The dozens of wards, offices, and kitchens,
serving the hundreds of patients and staff, generates considerable
infectious medical waste, mixed domestic and office waste, and kitchen
waste, which needs to be removed and disposed of daily. Some individual
wards implement waste separation, binning medical wastes separately from
general waste, however separation ends at the point of collection.
Queen's is not served by regular municipal waste collection. What
municipal collection that is done is ad-hoc, sporadic, and based on the
hospital's available financial resources. Rather, most hospital waste
(infectious and non-infectious) is gathered by grounds staff and burnt
at the hospital's incinerator, located at a central point within the
hospital campus[^1] [@kalina2020this]. However, the vast majority of the
deposited waste is not incinerated, as a new incinerator provided by the
Ministry of Health in late 2019, is only able to handle a small
percentage of the hospital's waste. Furthermore, the incinerator is
frequently inoperable, under repair, or without power[^2]. Rather, the
waste is openly burned, in a constantly smouldering pile to the side of
the incinerator building, sending up the aforementioned clouds of smoke,
which choke the grounds (Figure 1). Innumerable other, smaller, fires,
scattered across the QECH campus (Figure 2) contribute their part, as
grounds and maintenance staff habitually burn leaves and other garden
refuse. Furthermore, caregivers[^3] who reside separately on hospital
grounds, cook for themselves throughout the day and night using biomass
(charcoal or wood), and burn their own domestic waste. As a result, air
quality at the hospital is a constant source of discussion, with
patients, staff, and caregivers struggling to cope, when just breathing
is, as the quote used in the title of this article, 'unbearable.'

[^1]: Sharps are handled and disposed of separately.

[^2]: There was an older incinerator on the grounds which failed in
    2017. Between 2017 and 2021, when the new incinerator was
    commissioned, nothing was incinerated at all.

[^3]: Patients are required to bring a caregiver for in-patient
    hospitals stays, usually a family member, who assist with feeding,
    bathing, and all other non-medical services.

@@

-   **The Blantyre Malaria Project (BMP)**, established by Professors
    Terrie Taylor and Malcolm Molyneux, has carried out clinical
    research and patient care in the area of paediatric malaria,
    specifically cerebral malaria, for more than 30 years[^4]. Its
    research infrastructure includes an administrative team, an
    inpatient research ward, an MRI centre, and an outpatient research
    clinic in Ndirande township and in other districts outside of
    Blantyre.

@@

-   When a patient arrives at QECH, she or he will usually arrive with a
    guardian: someone to cook for them, buy medicine, do their laundry,
    and help them bathe. The **Guardian Shelter** is a gender-separated
    concrete floor shelter for sleeping along with a cooking pavilion
    and toilet/shower block are maintained by a the local NGO Chira
    fund. Though simple, the facility is secure and offers one of the
    only free/accessible toilet facilities for visitors on the QECH
    campus.

[^4]: BMP is credited with the development of the Blantyre Coma Score,
    which is used widely.

@@

The PM monitoring device centres around a Raspberry Pi 3 Model B
single-board computer complete with an operating system and storage
space. The added advantage of storage space is that the data are not
lost if the internet connection breaks down. Because the Pi can only
read digital signals, we needed to include an analog-to-digital (ADC)
converter. The amount of dust is measured by a Nova PM SDS011 High
Precision Laser sensor. This sensor measures particles at 2.5 and 10
micrometres in diameter in mg/m3. A list of the specific hardware and
software components can be found in Appendix B.

@@

Due to localised variation (air movement in the immediate area) and the
density of the observations, the plotted 5-minute data obscure
persistent trends. The full set of 5-minute data are plotted and
presented in Appendix A Figure A.1 (TODO). However, the number of measured values
that exceeded the hazardous limit for both parameters are summarised in
Table 2.

@@

# Appendix {.unnumbered}

@@

## Appendix A {.unnumbered}

```{r}

# Figure XA1. PM2.5 and PM10 values collected every 5 minutes over 10 months at 8 locations. 
# TODO: ##Lars can you please fix the x-axis labels so they aren’t squished

## Appendix B {.unnumbered}

### Hardware Components {.unnumbered}

1.  Raspberry Pi 3 Model B single-board computer technical
    specifications:
    -   Broadcom BCM2837 64bit Quad Core Processor powered Single Board
        Computer running at 1.2GHz 1GB RAM
    -   BCM43438 WiFi on board
    -   Bluetooth Low Energy (BLE) on board
    -   40pin extended GPIO
    -   4 x USB 2 ports 4 pole Stereo output and Composite video port
    -   Full size HDMI CSI camera port for connecting the Raspberry Pi
        camera
    -   DSI display port for connecting the Raspberry Pi touch screen
        display
    -   Micro SD port for loading your operating system and storing data
    -   Upgraded switched Micro USB power source (now supports up to 2.4
        Amps)
    -   Same form factor as the Pi 2 Model B, however the LEDs have
        changed position
    -   RPI3-MODB-16GB-NOOBS
2.  Nova PM SDS011 Sensor
    -   See Data Sheet attachment for technical specifications
3.  MCP3008 microchip
    -   analog-to-digital converter (ADC)
4.  Romoss Sense8+ 30000mAh QC Type-C
    -   Power bank battery
5.  Pipe
    -   clear rubber pipe attainable from any local hardware store
6.  Box
    -   any box that can fit all the components comfortbaly can be used.
        The box will need to have a whole drilled in it for the pipe.

### Software ‘components’ and data management {.unnumbered}

1. The Raspberry Pi operating system runs off of Rasbian software on a NOOBS scandisk. 
2. Air quality monitoring code is written in Python using the [aqi library](https://pypi.org/project/python-aqi/) (A library of algorithms to convert between AQI value and pollutant concentration). The code is stored on the [Open Data Durban GitHub account](https://github.com/opendatadurban/hospital_stations) and is free for all to access. To be able to collect data, the device needs to be connected to a reliable wifi network, the SSID and password need to be coded onto the device during step 2 in the building method.
3. Once the whole device has been put together and placed on site, to access that data we use a combination of propgrams:
    - [Putty](https://www.putty.org/) - this is an SSH client that we use to gain access to the Raspberry Pi from another computer. To be able to access the Raspberry Pi in question, the computer in use and the air quality monitoring device need to be on the same network for the connection to be successful. 
    - [Winscp](https://winscp.net/eng/download.php) - this is an FTP client that we use, in conjunction with Putty, to be able to transfer data files from the air quality monitoring device to one’s computer.
    - [Remote.it](https://app.remote.it/) - this is an application that supports remote SSH, so this is used when wanting to access data from a device that is not connected to the same network as the computer in use. Remote.it is used together with Putty and Winscp to access the air quality data remotely.

### Connection/building method {.unnumbered}

To build the sensor, we followed these steps:

1. Set up the Raspberry Pi with the Rasbian software using the NOOBS scandisk. (insert the scandisk into the Rasberry Pi) A comprehensive start up guide can be found here. 
2. Next we load the code onto the Raspberry Pi. The code can be found on the Open Data Durban Hospital Stations Repository Github. To use it you can follow the steps in the file called Initiate_WS.txt. Once all the steps have been completed, the Raspberry Pi is now a device that can be connected to a particulate matter monitoring sensor.  
3. Connect the hardware components
    - Connect the Raspberry Pi to the Nova PM Sensor using the break out board that comes with the Sensor. 
    - Connect the battery to the Raspberry Pi via the micro port
    - Attach the rubber pipe to the sensor 
    - Assemble all components inside the box, with the pipe coming out of the drilled hole
    - To prevent dirt entering the system, close off all openings between the pipe and the box with silicone putty
    - Secure box in an elevated position (not too close to the ground) and with the pipe facing the direction of the inflow of the air under investigation.
